---
title: "time Series"
author: "Sagar"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Time Series Analysis

> A time series is an ordered sequence of of equally spaced values over time. It is denoted by: Y = a + bX

* Components of a time series:
  + Trend
  + Seasonality
  + Cyclic
  + Random


### The Autoregressive(AR) Model

> The Autoregressive Model is denoted by: 
y~t~ = δ + φ~1~y~t-1~ + φ~2~y~t-2~ + … + φ~p~y~t-1~ + At
> A particular point in the series can be expressed as a linear combination of the prior p values.

### The Moving Average(MA) Model

> The Moving Average Model is denoted by:
y(t)=\mu +\varepsilon _{t}+\theta _{1}\varepsilon _{{t-1}}+\cdots +\theta _{q}\varepsilon _{{t-q}}\,
> In MA(q) model, the value of the time series is a linear combination of the error term or the residuals.

### Autoregressive Moving Average(ARMA) Model

> ARMA model is defined as a combination of the AR(p) and the MA(q) model. To apply a model to a time series, the series should be Stationary in nature.

* Criteria for a Stationary Series:
  + The mean is constant for the series
  + The variance of the series is finite
  + the covariance of the y~t~ and y~t+h~ depends only on the value of the lag
  
### **Steps in time series analysis**

#### Get the data 

```{r}
#using WDI data 
#install.packages("WDI")
library("WDI")
```

Using the GDP per capita from 1960 to 2016

```{r}
gdp <-  WDI(country = c("US"), indicator = c("NY.GDP.PCAP.CD"), start = 1960, end = 2016)
# looking at the data
head(gdp)
```

Changing the column name for the third column

```{r}
names(gdp) <-  c("iso2c", "coutry", "GDPPerCapita", "year")
head(gdp)
```


#### Organizing the data set in the ascending order:

```{r}
gdp <-  gdp[order(gdp$year), ]

```

#### Checking for stationary data assumption

```{r}
plot(gdp$GDPPerCapita~gdp$year)
```

Using ts function in R to transform the data to time series.

```{r}
tsgdp <-  ts(gdp$GDPPerCapita, start = min(gdp$year), end = max(gdp$year) )
```

Clearly there is a positive trend in the series.
Since the data insn't stationary using differencing to transform the current series to a stationary series.

> Using the forecast package for diffing

```{r}
#install.packages("forecast")
library("forecast")
```

```{r}
usGDP <-  auto.arima(tsgdp)
usGDP
```


We have a ARIMA(2,2,2) model with parameters 2,2,2
The first parameter is of the AR(q) model denoting, in order to analyze the gdp per capita of a year we need to consider the GDP of last two years.
The second parameter denotes the level of differencing, i.e. 2.
The third parameter is of MA(p) model that denotes, we need to consider the error term for the last two years.

y~t~ = φ~1~y~t-1~ + φ~2~y~t-2~ + $\alpha$~1~$\varepsilon$~t-1~ + $\alpha$~2~$\varepsilon$~t-2~


#### Calculating the coeeficients for the model

```{r}
coef(usGDP)
```

#### Predicting future values of the series for 10 years

```{r}
predictionResult = predict(usGDP, n.ahead=10, se.fit = T  )
predictionResult
```

#### Forecasting & plotting the results

```{r}
GDPForecast <-  forecast(object = usGDP, h =10)
plot(GDPForecast)
```




